---
title: "VAR_Cholesky"
output: html_document
date: "2025-07-11"
---

## Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tseries)
library(quantmod)
library(ggplot2)
library(vars)
library(data.table)
library(jtools)
library(dplyr)
library(zoo)
library(forecast)
library(tidyr)
library(dplyr)
library(tibble)
library(patchwork)
#install.packages("stargazer")
library(stargazer)
library(lpirfs)
library(tidyverse)
library(svars)

```

## Data Loading and Making a Data Frame

We first load the data and the transform it to turn it into a time-series object. We then create a "data" dataframe.

```{r dataframe, include=FALSE}

data <- read_excel("/Users/nejcperme/Desktop/Faks/Podiplomski - 2.letnik/Monetary Economics 2/SVAR/Empirical_exam.xlsx")

#Transform data
y <- ts(log(data$y)*100, start = c(2000, 1), freq = 12)  # output
p <- ts(log(data$hcpi)*100, start = c(2000, 1), freq = 12)  # consumer inflation
i <- ts(data$i, start = c(2000, 1), freq = 12)
carbon <- ts(data$carbon, start=c(2000,1), freq=12)
u <- ts(data$u, start=c(2000,1), freq=12)

# Plot data
data <- cbind(carbon, p, u,y, i)
data <- as.data.frame(data)
colnames(data) <- c("carbon","p","u", "y", "i")
plot.ts(data)
```

## VAR estimation

We estimate the VAR and estimate the SVAR - sturctural impulse responses with packages.

```{r VAR_packages, include = TRUE, warning=FALSE}
#--------------------- Manual Calculation -------------------------
#VAR estimation
var.est1 <- VAR(data, p = 3,type = "cons")

# Setting up the SVAR with Cholesky decomposition
a.mat <- diag(5)
diag(a.mat) <- NA
a.mat[2, 1] <- NA
a.mat[3, 1] <- NA
a.mat[3, 2] <- NA
a.mat[4, 1] <- NA
a.mat[4, 2] <- NA
a.mat[4, 3] <- NA
a.mat[5, 1] <- NA
a.mat[5, 2] <- NA
a.mat[5, 3] <- NA
a.mat[5, 4] <- NA
print(a.mat)

b.mat <- diag(5)
diag(b.mat) <- NA
print(b.mat)

svar.one <- SVAR(var.est1, Amat = a.mat, Bmat=b.mat, max.iter = 10000, 
                 hessian = TRUE)
#v svar ne mors dt identity matrixa za B recimo, pol sam matriko vn vrzs!!!
svar.one$A
svar.one$B

# Compute and store IRFs
par(mfrow = c(4, 1), mar = c(2.2, 2.2, 1, 1), cex = 0.6)
irf_y <- irf(svar.one, response = "p", impulse = "carbon", 
             n.ahead = 36, ortho = TRUE, boot = TRUE)
irf_y <- irf(svar.one, response = "y", impulse = "carbon", 
             n.ahead = 36, ortho = TRUE, boot = TRUE)
irf_i <- irf(svar.one, response = "i", impulse = "carbon", 
             n.ahead = 36, ortho = TRUE, boot = TRUE)
irf_u <- irf(svar.one, response = "u", impulse = "carbon", 
             n.ahead = 36, ortho = TRUE, boot = TRUE)
irf_carbon <- irf(svar.one, response = "carbon", impulse = "carbon", 
                  n.ahead = 36, ortho = TRUE, boot = TRUE)

plot(irf_y)
plot(irf_y)
plot(irf_i)
plot(irf_u)
plot(irf_carbon)

# Compute and store IRFs
two.int.pi <- irf(svar.one, response = "p", impulse = "i", 
                  n.ahead = 36, ortho = TRUE, boot = TRUE)
two.int.y <- irf(svar.one, response = "y", impulse = "i", 
                 n.ahead = 36, ortho = TRUE, boot = TRUE)
two.int.i <- irf(svar.one, response = "i", impulse = "i", 
                 n.ahead = 36, ortho = TRUE, boot = TRUE)
two.int.u <- irf(svar.one, response = "u", impulse = "i", 
                 n.ahead = 36, ortho = TRUE, boot = TRUE)
# Plot each IRF
plot(two.int.pi, main = "Response of p to monetary policy shock")
plot(two.int.y, main = "Response of y to monetary policy shock")
plot(two.int.i, main = "Response of i to monetary policy shock")
plot(two.int.u, main = "Response of u to monetary policy shock")
```

Note that ~ is a shorthand for function(.x). We estimate a VAR algebraically, using OLS. 
First method is the matrix way, whereas the second is done equation by equation. 
```{r VAR_manually, include = TRUE}
#---------------------- Getting the right dataframes ----------------------------
lagged_data <- data %>%
  mutate(across(everything(),
                .fns = list(
                  lag1 = ~lag(.x, 1),
                  lag2 = ~lag(.x, 2),
                  lag3 = ~lag(.x, 3)
                ),
                .names = "{.col}_{fn}"))
lagged_data$cons <- 1 
clean_frame <- na.omit(lagged_data)
Y <- as.matrix(clean_frame[, c("carbon","p", "u", "y", "i")])
regressors <- grep("_lag", names(clean_frame), value=TRUE)
X <- as.matrix(clean_frame[, c(regressors, "cons")])
#-------------------------------- OLS - matrix ------------------------------------------
B_hat <- solve(t(X)%*%X)%*%t(X)%*%Y #OLS matrix (X'X)^-1 X'Y
U_hat <- Y-X%*%B_hat #residual matrix
T <- nrow(X) #nobs
Sigma_u <- (t(U_hat) %*% U_hat)/ (T-ncol(X)) #variance covariance matrix of residuals
invXtX <- solve(t(X)%*% X)
vcov_B <- kronecker(Sigma_u, invXtX) #variance covariance matrix of regression coefficient
std_err <- sqrt(diag(vcov_B))
se_matrix <- matrix(std_err, nrow = nrow(B_hat), ncol = ncol(B_hat))
colnames(se_matrix) <- colnames(B_hat)
rownames(se_matrix) <- colnames(X)
results <- var.est1

# Extract standard errors for each equation
se_list <- lapply(results$varresult, function(eq_summary) {
  coef_table <- summary(eq_summary)$coefficients
  coef_table[, "Std. Error"]
})
coef <- var.est1$varresult$carbon$coefficients

#-------------------------------- OLS - eq-by-eq ------------------------------------------
carbon_reg <- lm(carbon ~ carbon_lag1 + carbon_lag2+carbon_lag3+p_lag1+p_lag2+p_lag3+u_lag1+u_lag2+u_lag3+y_lag1+y_lag2+y_lag3+i_lag1+i_lag2+i_lag3, data=lagged_data)
p_reg <- lm(p ~ carbon_lag1 + carbon_lag2+carbon_lag3+p_lag1+p_lag2+p_lag3+u_lag1+u_lag2+u_lag3+y_lag1+y_lag2+y_lag3+i_lag1+i_lag2+i_lag3, data=lagged_data)
y_reg <- lm(y ~ carbon_lag1 + carbon_lag2+carbon_lag3+p_lag1+p_lag2+p_lag3+u_lag1+u_lag2+u_lag3+y_lag1+y_lag2+y_lag3+i_lag1+i_lag2+i_lag3, data=lagged_data)
u_reg <- lm(u ~ carbon_lag1 + carbon_lag2+carbon_lag3+p_lag1+p_lag2+p_lag3+u_lag1+u_lag2+u_lag3+y_lag1+y_lag2+y_lag3+i_lag1+i_lag2+i_lag3, data=lagged_data)
i_reg <- lm(i ~ carbon_lag1 + carbon_lag2+carbon_lag3+p_lag1+p_lag2+p_lag3+u_lag1+u_lag2+u_lag3+y_lag1+y_lag2+y_lag3+i_lag1+i_lag2+i_lag3, data=lagged_data)
vcov_carbon_reg <- vcov(carbon_reg)
vcov_p_reg <- vcov(p_reg)
vcov_y_reg <- vcov(y_reg)
vcov_u_reg <- vcov(u_reg)
vcov_i_reg <- vcov(i_reg)
se_carbon_reg <- sqrt(diag(vcov_carbon_reg))
se_p_reg <- sqrt(diag(vcov_p_reg))
se_y_reg <- sqrt(diag(vcov_y_reg))
se_u_reg <- sqrt(diag(vcov_u_reg))
se_i_reg <- sqrt(diag(vcov_i_reg))
```


```{r pressure, echo=FALSE}
plot(pressure)
```

